---
title: "ANOVA"
output:
  html_document:
    css: ../lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
---

```{r echo = FALSE, message=FALSE}

knitr::opts_chunk$set(warning = FALSE, eval = TRUE, results = TRUE, fig.show = "hold", message = FALSE)
library(tidyverse)
library(hrbrthemes)
library(knitr)
library(kableExtra)
library(car)
```

### Introduction

In this lab, we will learn about Analysis of Variance (ANOVA). ANOVA is a type of linear model  that can be used to compare the means of two or more groups to determine if there are any significant differences between them. It is an extension of the t-test, which can be used to compare the means of two groups.

ANOVA is used when we have categorical independent variables (also known as factors) and a continuous dependent variable. It helps us understand if the means of the dependent variable differ significantly across different levels of the independent variable.

### The Data

Suppose we are interested in studying the effectiveness of three different study techniques (A, B, and C) on student performance. We randomly assign 60 students to these three study techniques, with 20 students in each group. After a week of studying, we measure their test scores. The data are as follows:
  

Suppose you are interested in studying the average height of students at your university. You randomly select 20 students and measure their height in inches. The data are as follows:

```{r}

set.seed(1)

study_data <- data.frame(
  study_technique = rep(c("A", "B", "C"), each = 20),
  test_scores = rnorm(60, mean = c(30,  60, 85), sd = c(5))
)

kable(
study_data %>% head,
    "html") %>% 
  kable_styling(font_size = 10,
                "striped") %>% 
  scroll_box(width="100%")

```

### Setting up the hypothesis test

Setting up the Hypothesis Test
To perform the ANOVA, we need to specify the null hypothesis and alternative hypothesis. The null hypothesis assumes that the means of all groups are equal, while the alternative hypothesis assumes that at least one group's mean is different.

* Null hypothesis: The means of study techniques A, B, and C are equal.
* Alternative hypothesis: At least one study technique's mean is different.

```{r}

study_data %>% 
  ggplot(aes(x=study_technique,y=test_scores)) +
  geom_boxplot(aes(fill=study_technique), alpha=0.6) +
  theme_ipsum() +
  theme(
    legend.position = 'None'
  )

```

### Checking the Assumptions

Before conducting the ANOVA, we need to check the assumptions. The assumptions for ANOVA are:

* Independence: The observations within each group should be independent of each other.
* Normality: The data within each group should be approximately normally distributed.
* Homogeneity of variances (homoscedasticity): The variability (variance) of the data within each group should be approximately equal.

We can check the normality assumption by examining the histograms and normal probability plots for each group.

```{r}
ggplot(study_data, aes(x = test_scores)) +
  geom_histogram(aes(y = ..density..), bins = 8, color = "darkgrey", fill = "aquamarine", alpha=0.6) +
  geom_density(color = "darkblue", lwd = 0.9, linetype='dotted') +
  facet_wrap(~ study_technique, ncol = 3, scales = "free") +
  labs(title = "Histogram of Test Scores by Study Technique", x = "Test Scores", y = "Density") +
  theme_ipsum()

```

```{r}

ggplot(study_data, aes(sample = test_scores)) +
  stat_qq(color = "darkblue") +
  stat_qq_line(color = "darkgrey") +
  facet_wrap(~study_technique, ncol = 3) +
  labs(title = "Normal Probability Plot of Test Scores by Study Technique", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_ipsum()

```

Based on the histograms and normal probability plots, the data within each group appear to be approximately normally distributed.
  
To check the homogeneity of variances assumption, we can perform Levene's test, which tests if the variances of the groups are equal.
  
```{r}  

leveneTest(test_scores ~ study_technique, data = study_data)

```

The p-value for Levene's test is greater than 0.05, indicating that we do not have enough evidence to reject the null hypothesis of equal variances. Therefore, we can assume the homogeneity of variances assumption holds.
  
### Performing ANOVA

```{r}

anova_result <- aov(test_scores ~ study_technique, data = study_data)
summary(anova_result)

```
  
### Interpreting the Results
  
The ANOVA test does not tell us which groups are significantly different from each other but rather if there is a significant difference between any of the groups. The ANOVA table shows the results of the analysis. The p-value for the study_technique variable is more than 0.05, indicating that there is not enough evidence to reject the null hypothesis of at least one mean of the samples being different from the others.